A Comparison of Parallel Graph Processing Implementations

### 2017-05-16 papers/publication/paper-pollard.tex
Added scale 23. Turning in again.

\subsection{Scalability}
Figures~\ref{fig:bfs-speedup} and~\ref{fig:bfs-efficiency} illustrate the parallel strong efficiency of the BFS algorithm in different packages on a Kronecker graph of scale $20$ . The parallel speedup shown in Fig.~\ref{fig:bfs-speedup} is computed as $T_1 / T_n$ where $T_1$ is the sequential time and $T_n$ is the execution time on $n$ threads. 
\begin{figure}[htb]
	\centering
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup20.pdf}
		\caption{Speedup of BFS for a scale-20 graph, black solid line represents ideal speedup. Both axes are logarithmic with an exception at 72 threads for readability. }
	\label{fig:bfs-speedup}
\end{figure}

Figure~\ref{fig:bfs-efficiency} shows the parallel efficiency,  $T_1 / (nT_n)$ for different implementations of BFS. Ideal efficiency is defined as $T_n = T_1/n$ and is the horizontal line near the top of Fig.~\ref{fig:bfs-efficiency}. 
\begin{figure}[htb]
	\centering
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss20.pdf}
	\caption{Parallel efficiency for a scale-20 graph. GraphBIG dips below 1 because it is slower for $2$ threads than for 1. $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time with $n$ threads.}
	\label{fig:bfs-efficiency}
\end{figure}

These plots show generally poor scaling for this size problem, a challenge facing most parallel graph algorithms. In addition, $2^{20}$ vertices is a relatively small graph by today's standards and thus library designers focus on scalability for larger graphs. Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine. GCC version 4.8.5 supports OpenMP version 3.1 but the most recent release of OpenMP is 4.5.

@@

Overall, the GAP Benchmark Suite was the best-performing system across all chosen datasets, although not the most scalable for the graphs used in this study (at most $2^{22}$ vertices). In addition, GAP is the most recent project and also the most robust in terms of the input graphs accepted. Thus, we recommend graph processing algorithm designers compare their BFS implementations to the Graph500 for BFS, GAP Benchmark Suite for smaller problem sizes or a low number of threads (at most 32), and GraphMat for larger problems with more threads. This is despite the fact that GraphBIG has the best scalability, because even at more than 16 threads the runtime for GraphBIG is still at least one order of magnitude slower than GAP or GraphMat for all sampled algorithms. Last, but not least, one must consider cost and portability: GraphMat requires the Intel compiler collection, which is not free or universally available for all platforms.

>>>>

\subsection{Scalability}
Figures~\ref{fig:bfs-speedup} and~\ref{fig:bfs-efficiency} illustrate the parallel strong efficiency of the BFS algorithm in different packages on a Kronecker graph of scale $23$. The parallel speedup shown in Fig.~\ref{fig:bfs-speedup} is computed as $T_1 / T_n$ where $T_1$ is the sequential time and $T_n$ is the execution time on $n$ threads. Because of timing considerations, only four trials were run for these experiments.
\begin{figure}[htb]
	\centering
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup23.pdf}
		\caption{Speedup of BFS for a scale-20 graph, black solid line represents ideal speedup. Both axes are logarithmic with an exception at 72 threads for readability. }
	\label{fig:bfs-speedup}
\end{figure}

Figure~\ref{fig:bfs-efficiency} shows the parallel efficiency,  $T_1 / (nT_n)$ for different implementations of BFS. Ideal efficiency is defined as $T_n = T_1/n$ and is the horizontal line near the top of Fig.~\ref{fig:bfs-efficiency}. 
\begin{figure}[htb]
	\centering
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss23.pdf}
	\caption{Parallel efficiency for a scale-23 graph. Graph500 dips below 1 because it is slower for $2$ threads than for 1. $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time with $n$ threads.}
	\label{fig:bfs-efficiency}
\end{figure}

These plots show generally poor scaling for this size problem, a challenge facing most parallel graph algorithms. In addition, $2^{23}$ vertices is a relatively small graph by today's standards and thus library designers focus on scalability for larger graphs. Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine. GCC version 4.8.5 supports OpenMP version 3.1 but the most recent release of OpenMP is 4.5.

Overall, GAP is the most scalable with GraphMat close behind for larger threads and even slightly beating GAP at 72 threads. While care was taken to ensure there was no other load on the system when the experiments were performed, the efficiency for two threads is surprisingly low for the Graph500. Because the Graph500 spends a shorter amount of time executing in general (there is no file I/O performs multiple BFS passes one right after another), it is more sensitive to spikes in CPU usage.

@@

Overall, the GAP Benchmark Suite is the best-performing system across all chosen datasets and in general the most scalable for the graphs used in this study (at most $2^{23}$ vertices). In addition, GAP is the most recent project and also the most robust in terms of the input graphs accepted.  Powergraph and Graphmat are designed for larger graphs (it is possible to run in distributed mode for both) so the overhead of these frameworks may dominate for smaller problem sizes. Two more potentially important considerations are cost and portability: GraphMat requires the Intel compiler collection which may be cost-prohibitive for some users. Despite the GAP Benchmark Suite being the overall best performing benchmark, it was not the best performing in every case so we unsurprisingly recommend the more comprehensive comparison using all five benchmarks offered by \emph{easy-parallel-\mbox{graph-\textasteriskcentered}}.
